---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r,echo=TRUE}
# setwd("~/Google Drive/GitHub/RepData_PeerAssessment1")
# setwd("C:/Users/Ben/Google Drive/GitHub/RepData_PeerAssessment1")
# setwd("C:/Users/Harry Samsung Laptop/Downloads/ben.liu's work/GitHub/RepData_PeerAssessment1")

unzip("activity.zip")
# list.files()
data=read.csv("activity.csv",colClass=c("integer","Date","factor"))

library(ggplot2)
library(lattice)

```

## What is mean total number of steps taken per day?
### histogram of the total number of steps taken each day
```{r,echo=TRUE}
# remove NA in steps column
df=data[!is.na(data$steps),]

dailysum=aggregate(steps~date,data=df,sum)
qplot(steps, data=dailysum,main = "Histogram of total number of steps taken each day")
```

### the mean total number of steps taken per day
```{r,echo=TRUE}
dailymean=aggregate(steps~date,data=df,mean)
dailymean
```

### the median total number of steps taken per day
```{r,echo=TRUE}
dailymedian=aggregate(steps~date,data=df,median)
dailymedian
```

## What is the average daily activity pattern?
### time series plot of the 5-minute interval averaged across all days
```{r,echo=TRUE}
# allocate order number for intervals; 17568/288=61
# data$int=rep(1:288,61)
data$interval=as.numeric(as.character(data$interval))

# aggregate steps with same interval/time of the day; NA will be removed
fivemin=aggregate(steps~interval,data=data,mean,na.rm=TRUE)

# total number of steps taken each day
ggplot(fivemin,aes(x=interval,y=steps))+ geom_line() + ggtitle("time series plot of the 5-minute interval averaged across all days") + labs(x="5-min interval",y="Average steps") 

```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
The 5-minute interval containing the maximum number of steps is:
```{r,echo=TRUE}
fivemin$interval[fivemin$steps %in% max(fivemin$steps)]
```


## Imputing missing values
### Calculate and report the total number of missing values in the dataset
Total number of missing values is:
```{r,echo=TRUE}
nrow(data[is.na(data$steps),])
```

### Devise a strategy for filling in all of the missing values in the dataset
Use average (mean) 5-minute interval to fill in missing values.

### Create a new dataset with the missing data filled in
```{r,echo=TRUE}
df1=data[!is.na(data$steps),]
df2=data[is.na(data$steps),]
for(i in 1:nrow(df2)){
  df2$steps[i]=fivemin$steps[fivemin$interval %in% df2$interval[i]]
}

# create new dataset
newdata=rbind(df1,df2)

# rm(df1,df2)

```

### Make a histogram of the total number of steps taken each day
```{r,echo=TRUE}
dailysumnew=aggregate(steps~date,data=newdata,sum)
qplot(steps, data=dailysumnew,main = "Histogram of total number of steps taken each day")
```

### Calculate and report the mean and median total number of steps taken per day. 

#### new mean total number of steps taken per day
```{r,echo=TRUE}
dailymeannew=aggregate(steps~date,data=newdata,mean)
dailymeannew
```

#### new median total number of steps taken per day
```{r,echo=TRUE}
dailymediannew=aggregate(steps~date,data=newdata,median)
dailymediannew
```
### Compare with estimates from first part of assignment

+ Difference in mean of total number of steps taken per day:
```{r,echo=TRUE}
mean(dailymean$steps)-mean(dailymeannew$steps)
```

+ Difference in median of total number of steps taken per day:
```{r,echo=TRUE}
mean(dailymedian$steps)-mean(dailymediannew$steps)
```

### Impact of imputing missing data on the estimates of the total daily number of steps
On average, the mean of daily total steps doesn't change, while the median increases. 


## Differences in activity patterns between weekdays and weekends

### Create a new factor variable in the dataset with two levels - "weekday" and "weekend"

```{r,echo=TRUE}
# drop weekdays() function 
library(timeDate)
newdata$day[isWeekday(newdata$date,w=1:5)]="weekday"
newdata$day[!isWeekday(newdata$date,w=1:5)]="weekend"

```


### Make a panel plot containing a time series plot of the 5-minute interval and the average number of steps taken across all weekday days or weekend days

```{r,echo=TRUE}
dailymeanwk=aggregate(steps~day+interval,data=newdata,mean)

# library(lattice)
with(dailymeanwk,xyplot(steps~interval|day,type="l",layout=c(1,2),xlab="5-minute interval",ylab="Average steps"))

```